---
title: AP Statistics Chapter 5 Assignment
subtitle: Dungeons & Dragons (5th Edition) Monster Armor Classes
author: Caleb Smith
output: pdf_document
urlcolor: blue
---

# About

*Dungeons & Dragons* is a popular role-playing tabletop game, within which players are often pitted in combat against a variety of monsters. The 5th Edition *Monster Manual* is the definitive guidebook to the monsters found in the game (it is officially published by the creators of D&D), and contains 761 such creatures. Each monster has a number of game statistics - one such is the creature's *Armor Class*, which determines how difficult it is for a player to deal damage to the monster. In this report, I've gathered and analyzed the Armor Classes of each creature in the Monster Manual. 

# 1. Dataset

Below is the data set for each of the monsters in the \textit{Monster Manual} - values are listed in alphabetical order by the monster's name.

```{r}
ac <- read.csv("data.csv", header=TRUE)
ac$AC
```

# 2. Data Source

I found the data online in the form of Markdown documents, one per monster, which can be found [here](https://github.com/karagenit/bestiary/tree/master/_posts). Pages were parsed for data with [this script](https://github.com/karagenit/dnd-ac/blob/master/script.rb), which got the data into a basic CSV format (both as a number of counts per AC and as the raw AC values, the latter of which I used for this document). I then used [this script](https://github.com/karagenit/dnd-stats/blob/master/colsplit.sh) to split the columns into rows and manually added the AC header. 

Results of the script can be found [on this Google Doc](https://docs.google.com/spreadsheets/d/1GczVGTTNPxHVgNRFSQclzjwKyO30wLdCg6uDStq-raQ/edit?usp=sharing), or in [this CSV file](https://github.com/karagenit/dnd-stats/blob/master/data.csv).

A little about this document: I was originally planning on using LaTeX with latexpdf to generate a nice PDF report. However, I recently learned about the R statisics language, and decided to format this document with RMarkdown (the source of which can be found [here](https://github.com/karagenit/dnd-stats/blob/master/dnd.rmd)) and build it through RStudio (which uses a combination of pandoc + texlive). 

Originally I had to split the data with [this script](https://github.com/karagenit/csv-split/blob/master/script.rb) and then tabulate it for LaTeX with [this](https://github.com/karagenit/csv-latex-table/blob/master/script.sh), but R makes it easy to import and manipulate the data straight from CSV files.

# 3. Data Distribution & Normality

```{r}
hist(ac$AC,
     main = "Histogram of Armor Classes",
     xlab = "Armor Class",
     xlim = c(5,25),
     breaks = 20)
```

```{r}
boxplot(ac$AC,
        main="Boxplot of Armor Classes",
        horizontal = TRUE)
```







